Understanding Parameterized Quantum Circuit Learning for Quantum Chemical Applications

### 2025-03-31 manuscript/main.tex
Mon 31 Mar 2025 17:06:36 EDT

An important aspect of ML in chemistry is the choice of molecular representation, or how the molecule is represented in the machine learning models.\cite{jones_molecular_2023}
In general, molecular representations can be partitioned into three groups, graph-, topology-, and physics-based representations, all of which were explored to provide a comprehensive overview of how they perform using classical models as a baseline.
During the preprocessing stage, before converting to the XYZ coordinates into our molecular representation of choice, the set of hypothetical structures was reduced to 2436 molecules due to issues with valency exceptions when converting XYZ coordinates into RDKit mol objects.
Using RDKit\cite{noauthor_rdkit_nodate} we examined three commonly applied graph-based molecular preresentations, Molecular ACCess Systems (MACCS)\cite{durant_reoptimization_2002}, Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}, and RDKit fingerprints.
All three of these methods use traversals of the molecular graphs to encode various structural details into bit vectors.
Lastly, we explore both topology- and physics-based molecular representations, both of which encode the three-dimensional structure of molecules in various, unique ways.
Persistent images (PIs) are a topology-based fingerprint that uses persistence homology to encode topological information of three-dimensional molecular structures into fixed-dimension images.\cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022} 
We use the implementation from Townsend \textit{et al.}\cite{townsend_representation_2020}, which uses the Ripser Python package to generate PIs.\cite{tralie_ripserpy_2018}
Lastly, we explore two physics-based representations, Coulomb matrices (CMs) \cite{rupp_fast_2012} and smooth overlap of atomic positions (SOAPs), that were generated using DScribe.\cite{de_comparing_2016}
Both methods encode physical information regarding the atomic environments of each molecule, where CMs encode the Coulomb repulsion between atoms using nuclear charges, while SOAPs encode the similarities between atomic neighborhoods using kernels.
Due to the computational cost associated with computing the regularized entropy match (REMatch) kernel over the set of 2436 molecules with the SOAPs representation, we excluded SOAPs from the overall discussion.
For each of the molecular representations, we analyzed two different methods for representing the components of the BSE reaction.
The first represents the feature vector as the products subtracted from the reactants, denoted as \textit{sub} (e.g., $\text{CM}_{\mathit{sub}} = (\text{CM}_{\text{A}^{.}} + \text{CM}_{\text{B}^{.}}) - \text{CM}_{\text{A-B}}$), similar to the method used in Ref. \cite{garcia-andrade_barrier_2023}, and the second represents the feature vector using only the reactant (A-B), denoted as \textit{AB} (e.g., $\text{CM}_{\mathit{AB}} =  \text{CM}_{\text{A-B}}$).

>>>>

An important aspect of ML in chemistry is the choice of molecular representation, or how the molecule is represented in the machine learning models.\cite{jones_molecular_2023}
In general, molecular representations can be partitioned into three groups, graph-, topology-, and physics-based representations, all of which were explored to provide a comprehensive overview of how they perform using classical models as a baseline.
During the preprocessing stage, before converting to the XYZ coordinates into our molecular representation of choice, the set of hypothetical structures was reduced to 2436 molecules due to issues with valency exceptions when converting XYZ coordinates into RDKit mol objects.
Using RDKit\cite{noauthor_rdkit_nodate} we examined three commonly applied graph-based molecular preresentations, Molecular ACCess Systems (MACCS)\cite{durant_reoptimization_2002}, Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}, and RDKit fingerprints.
All three of these methods use traversals of the molecular graphs to encode various structural details into bit vectors.
Lastly, we explore both topology- and physics-based molecular representations, both of which encode the three-dimensional structure of molecules in various, unique ways.
Persistent images (PIs) are a topology-based fingerprint that uses persistence homology to encode topological information of three-dimensional molecular structures into fixed-dimension images.\cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022} 
We use the implementation from Townsend \textit{et al.}\cite{townsend_representation_2020}, which uses the Ripser Python package to generate PIs.\cite{tralie_ripserpy_2018}
Lastly, we explore two physics-based representations, Coulomb matrices (CMs) \cite{rupp_fast_2012} and smooth overlap of atomic positions (SOAPs), that were generated using DScribe.\cite{de_comparing_2016}
Both methods encode physical information regarding the atomic environments of each molecule, where CMs encode the Coulomb repulsion between atoms using nuclear charges, while SOAPs encode the similarities between atomic neighborhoods using kernels.
Due to the computational cost associated with computing the regularized entropy match (REMatch) kernel over the set of 2436 molecules with the SOAPs representation, we excluded SOAPs from the overall discussion.
For each of the molecular representations, we analyzed two different methods for representing the components of the BSE reaction.
The first represents the feature vector as the products subtracted from the reactants, denoted as \textit{sub} (e.g., $\mathbf{X}_{\mathit{sub}} = (\mathbf{X}_{\text{A}^{.}} + \mathbf{X}_{\text{B}^{.}}) - \mathbf{X}_{\text{A-B}}$), similar to the method used in Ref. \cite{garcia-andrade_barrier_2023}, and the second represents the feature vector using only the reactant (A-B), denoted as \textit{AB} (e.g., $\mathbf{X}_{\mathit{AB}} =  \mathbf{X}_{\text{A-B}}$).
